Comment on Hodgkins et al. (2018): Predicting Absolute Holocellulose and Klason Lignin Contents for Peat Remains Challenging

### 2022-07-25 analysis/peer_review_manuscript_1/1/reply-reviewer_1-1.Rmd

---
title: "Reply to Comments of Reviewer 1"
author: "Henning Teickner, Klaus-Holger Knorr"
date: "2022-07-24"
output: 
  html_document: default
  pdf_document:
    latex_engine: xelatex
    fig_caption: true
    keep_tex: true
bibliography: ../../paper/references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reviewer comments have prefix "Q" and our answers prefix "A". "old" in combination with line numbers refers to lines in the previous version of the manuscript. "new" in combination with line numbers refers to lines in the updated version of the manuscript.

## General comments

> **Q1:** "This is an excellent manuscript based on an excellent rationale with sound research
questions. It is very well structured and very well written. Building on previous data and
models, Teickner & Knorr provide detailed explanations of the steps followed to evaluate
the models, describing the limitations (representative of the training data sets, validation
of the models, availability of key data of the models’ output, biases and uncertainties,
etc.) and potential improvements. This research also shows the key importance of a
detailed analysis of the models’ residuals as a way to identify model’s deficiencies. It is
also a key finding that OM composition of mineral soils can also be accurately modeled.
This research also shows the potential of MIR for SOM characterization and modelling and
the importance of making data and code available for further improvements.

> Regarding the general issue of the preselection of peaks before modelling, one possible
alternative is to perform PCA on all detected peaks to reduce the dimension and extract
the most relevant spectral signals. PCA is usually efficient in allocating into different
components the overlapping effects of more than one compound on a given spectral
band/region."

**A1:** Thank you for this encouraging comment. It is true that there are many more modeling approaches which probably would have a similar predictive accuracy as the approach we used here (especially since the sample size is small). For example, reviewer 2 suggested partial least squares regression (PLSR) as yet another alternative approach and there are many more approaches one could have tried (e.g. supervised PCR, iterative supervised PCR, interval PLSR, moving window interval PLS, etc; see e.g. @Xiaobo.2010).  
We have the impression that PLSR is more commonly used than PCR since PLSR is more efficient in identifying latent variables related to the target variable being predicted (since it maximizes also the covariance of the computed latent variables to the target variable, whereas PCR only maximizes the variance of the computed latent variables, which comprise only the spectral data). That said, we think that it is good that different tools exist and, depending on the problem, one or the other approach may result in a better predictive accuracy. PCR (or a different dimension reduction approach) may be particularly useful if more data becomes available so that the computational expenses of our approach get too heavy [@Piironen.2020].  
One advantage of the approach we used in comparison to dimension reduction approaches (e.g. PCR, PLSR), which summarize the original predictors into latent variables, is that coefficients for individual predictors are estimated more independently, since no latent variables are computed as summary of all predictors. Therefore, multiple regression with regularizing priors has the advantage of facilitating model interpretation (see e.g. Fig. 3).  
We think that it is good to at least briefly mention these aspects to provide readers a better orientation. To this end, we added the following paragraph (old: l. 175; new: l. 185): "An alternative, popular, approach to approach 2 would be dimension reduction, for example via partial least squares regression, principal component regression, or variants of these [@Xiaobo.2010]. In general, there are many alternative approaches which could be tested to use more information contained within the spectra than the original models, and many of these probably would result in similar predictive performances as approach 2 (regularization), especially when sample sizes are small [@Xiaobo.2010; @Teickner.2022]. An advantage of regularization is that model coefficients are estimated more independently than in dimension reduction approaches, which makes it more straightforward to interpret model coefficients. The key is that the approaches we chose are suitable to analyze our research questions."

--------------------------------------------

## Specific comments

> **Q2:** "L277-279. Positive and negative coefficients for peaks at 1150 and 1270 cm-1 (aromatic CH bending). This may be due to they reflecting different lignin structures: 1270 cm-1 surely corresponds to guaiacol (G) moieties and 1150 cm-1 could correspond to syringol (S) moieties (although S most representative band is at 1310 cm-1). Both differ in content in the source peat vegetation and also show differences in resistance to degradation in peat depending on oxygen availability (see for example, Schellekens et al. 2012 Soil Biology and Biochemistry 53, 32-42). G lignin is less prone to degradation than S lignin and the G/S ratio of fen peat (more decomposed) is larger than that of bog peat (less decomposed) (see for example Martínez Cortizas et al., 2021 Boreas 50, 1161-1178)."

**A2:** Thank you for this information, this is indeed interesting. Building on this: The G/S ratio is also different in the training dataset sample types. Softwood samples and needle samples have a higher G content than hardwood samples and leaves and grasses [@DelaCruz.2016]. They also have a (slightly) lower Klason lignin content than hardwood samples and leaves and grasses (supplementary Fig. S4).  
This indicates that since the bin at 1150 cm$^{-1}$ is caused to some extent by S, and S has a larger abundance in samples with smaller (than average) Klason lignin contents, this may explain why the coefficient for this bin is negative. Likewise, the bin at 1270 cm$^{-1}$ is caused to some extent by G, G has a larger abundance in samples with larger (than average) Klason lignin content, and this may explain why the coefficient for this bin is positive.  
To consider this hypothesis in the main text, we changed the text to (old: l. 277 to 279; new: l. 288): "A plausible explanation for the negative sign of the coefficient for the 1150 cm$^{-1}$ bin is that absorbance in this range is partly caused by syringyl units (S) [@Kubo.2005] and that training samples with higher S content have smaller Klason lignin contents (supplementary Fig. 4, @DelaCruz.2016), making this bin indicative of smaller Klason lignin contents. Likewise, a plausible explanation for the positive sign of the coefficient for the 1270 cm$^{-1}$ bin is that absorbance in this range is caused predominantly by guaiacyl units (G) [@Kubo.2005] and that training samples with higher G content have smaller Klason lignin contents (supplementary Fig. 4, @DelaCruz.2016), making this bin indicative of larger Klason lignin contents."

--------------------------------------------

> **Q3:** "L.424 I find this “implication” to be quite important and other ways of spectral normalization should be tested in the future."

**A3** We completely agree with this statement. We currently are not aware of a solution to this problem, other than to specify precisely under which conditions a prediction model produces erroneous predictions.

--------------------------------------------

```{r child="reply-additional.Rmd"}
```

## References

### 2022-07-25 analysis/paper/001-paper-main.Rmd
Address reviewer comments to manuscript 1.

    To understand global soil organic matter (SOM) chemistry and its dynamics, we need tools to efficiently quantify SOM properties, for example prediction models using mid infrared spectra. However, the advantages of such models rely on their validity and accuracy. Recently, \cite{Hodgkins.2018} developed models to quantitatively predict organic matter holocellulose and Klason lignin contents, two indicators of SOM stability and major fractions of organic matter. The models may have the potential to understand large-scale SOM gradients and have been used in various studies.  

@@

    The major limitation we identified is that the original training data are not representative for a range of diverse peat samples. This causes both biased estimates and extrapolation uncertainty under the original models. In addition, the original models can in practice produce unrealistic predictions (negative values or values $>100$ mass-%). Our improved models partly reduce the observed bias, have a better predictive performance for the training data, and avoid such unrealistic predictions. Finally, we provide a proof-of-concept that holocellulose contents can also be predicted for mineral-rich samples.  

@@

OM derived from plants and SOM is often characterized by step-wise chemical fractionation into holocellulose (acid soluble) --- a proxy for polysaccharides --- and lignin (acid insoluble) --- a proxy for aromatics --- [@DelaCruz.2016; @Elle.2019] and these variables are often important indicators for the chemical quality of OM in studies and models analyzing decomposition of SOM [@Leifeld.2012; @Biester.2014; @Worrall.2017; @Hodgkins.2018; @Bengtsson.2018; @Agren.1996a; @Bauer.2004; @Shipley.2021]. If these fractions can be predicted from MIRS, it may be possible to understand decomposition across larger scales and at higher spatial resolution.  
Several models to predict holocellulose and lignin contents in different OM types have been developed (some based on near infrared spectra) [summarized for lignin by @Elle.2019; for holocellulose see e.g. @Peltre.2011; @Sun.2011]. Most of these models consider only wood, material from few species, or specific vegetation organs [@Elle.2019]. Another problem is that most of these models cannot be easily reproduced and used since the raw data and code have not been published.  
Recently, @Hodgkins.2018 developed models to predict Klason lignin and holocellulose contents from attenuated total reflectance MIRS. The models have several advantages over existing approaches. The training data comprises several different OM types, such as paper products (cardboard, office paper, magazines, newspaper), leaves from diverse species (trees and graminoids), and wood samples [@DelaCruz.2016; @Hodgkins.2018]. Moreover, both raw data and model code are publicly available [@Hodgkins.2018; @Teickner.2020a]. For this reason, the models are particularly suitable for application in other studies [@Teickner.2019; @Moore.2019; @Harris.2020; @Cong.2020; @Cong.2022], most recently in @Verbeke.2022 and @Baysinger.2022, and for future developments, e.g. by including additional data.   
A major problem is that neither @Hodgkins.2018, nor a later study using the models provided a thorough model validation. To compute the models, the authors developed a procedure to extract peaks which are indicative for specific OM fractions, such as holocellulose and Klason lignin (Fig. \@ref(fig:hodgkins-peaks-sample-fig01)), from the MIRS. The peaks are baseline corrected, their maximum height computed and divided by the sum of absorbance values in the spectra. These values are used to calibrate the prediction models. As model validation, only the linearity between the target variables and predictors was assessed (supplementary material to @Hodgkins.2018), but not for example how predicted values match measured values. Moreover, no information has been provided whether the training data is representative for SOM, particularly the peat samples analyzed in the study. Furthermore, it is unclear if pre-selecting peaks may reduce the predictive accuracy of the models. Many of these concerns have also been raised by an anonymous reviewer of the paper (supplementary information to @Hodgkins.2018). Thus, an important research gap to fill is to validate the models and to provide concrete strategies for further improvements.

@@

Our goals are to identify limitations of the original models and to give concrete recommendations for improvements. Moreover, we use the original data from @Hodgkins.2018 to provide improvements on the original models where possible. To this end, we conducted an exploratory analysis using the peat and peat-forming vegetation data provided by @Hodgkins.2018. Our exploratory analysis was guided by the following research questions:

1. _Is the normal distribution reasonable to predict holocellulose and Klason lignin contents for peat samples?_ Klason lignin and holocellulose contents as actually measured for the training data set in @Hodgkins.2018 cannot be negative or larger than 100 mass-%. For this reason, predictions and prediction uncertainties of the models should be in the interval [0, 100] mass-%. Yet, the original models assume a normal distribution which does not entail any lower or upper bound for predicted values. This can potentially result in unrealistic predictions. However, if no unrealistic predictions occur for "representative" peat samples, using a model with normal distribution can be justifiable in practice due the central limit theorem. In contrast, if unrealistic predictions occur, it is more reasonable to use a distribution which assumptions are in accordance with knowledge on the data generating process. For example, the beta distribution assumes a lower and upper bound which can be mapped to the interval [0, 100] mass-%. For this reason, we were interested if and under which conditions the original models may produce unrealistic predictions for the peat and vegetation samples.

@@

3. _Is the prediction domain of the training data representative for peat samples?_ If a model extrapolates outside the range of predictor variable values in the training data set, it is unclear if the predictions are valid. An advantage of the independent reference sample set used by @Hodgkins.2018 is that the model is applicable on various OM types and e.g. no problem-specific correlations are used by the models to make predictions [@Hodgkins.2018]. For example, in peat, holocellulose content is controlled by decomposition, and other structures also controlled by decomposition might therefore provide information on holocellulose contents [@Leifeld.2012; @Biester.2014]. But this does not have to be the case for other samples, e.g. peat forming vegetation. An advantage is therefore the generality of the training data set @Hodgkins.2018 used. However, since it does not include peat samples, it is questionable if the prediction domain formed by the training samples covers peat samples and peat forming vegetation in general, and this can result in overlooking prediction failures due to extrapolation. For this reason, we were interested if the training data (its predictions domain, @Wadoux.2021) is representative for peat and peat forming vegetation.

4. _Do predictions of the improved models differ and if so, why?_ Computing modified models enables us to compare their predictions to those of the original models using the same training data, and to the actually measured contents. Model comparison often reveals detailed insights into the mechanics of a model and may provide hints to which model makes more correct predictions. We therefore were interested in analyzing differences between the models computed, what factors may cause differences in predictions, and if there are indications which models make more correct predictions.

5. _Can holocellulose contents also be predicted for samples with mineral interferences?_ During model calibration, @Hodgkins.2018 omitted some samples. For prediction of holocellulose, old magazine samples were omitted because the `carb` peak suffered from mineral interference from the samples' clay coatings [@Hodgkins.2018]. This decision is sensible, but since there are many organic soils with mineral admixtures --- for example due to volcanic ash, in base layers, or due to cryoturbation) [e.g. @Broder.2012; @Bockheim.2007; @Koven.2009; @Loisel.2014] --- we think that it is useful to have a model to predict holocellulose contents also for mineral rich samples. We therefore wanted to investigate if a suitable model can be computed by including the previously omitted old magazine samples during calibration.

@@

We conducted a series of statistical analyzes to query each of our research questions listed in the introduction. During this, we also computed improved models using the same training data that was used to compute the original models. Since we used these improved models to uncover and analyze the limitations of the original models, we first describe steps how we improved the original models and then how we analyzed the limitations.  
We use the data and models collected and developed by @Hodgkins.2018. The training data used to calibrate the original models consist of various organic materials and comprise measured values for holocellulose and Klason lignin [@DelaCruz.2016]. The data are available via @Hodgkins.2018. In addition, we used a data set consisting of MIRS for `r length(unique(d_peat$core_label))` peat cores (`r nrow(d_peat)` samples in total) and `r nrow(d_veg)` vegetation samples collected at the same sites which are also provided by @Hodgkins.2018. We used these peat and vegetation data to evaluate the performance of the original and improved models for peat and peat forming vegetation in general. We assume that the peat samples are representative for a wide range of SOM because they cover a large latitudinal range and a wide range of degrees of decomposition, and were formed by different vegetation [@Hodgkins.2018].  
We computed all models as Bayesian models. This allowed us to consider parameter uncertainty in predictions and to compute models with good predictive performance [@Raftery.2003; @Piironen.2017]. For this reason, we also recomputed the original models as Bayesian models with weakly informative priors to facilitate model comparison using the same metrics and tools. Whenever we refer to "original model", we mean these Bayesian translations, and otherwise refer to "original non-Bayesian models". To check that we did not introduce any of the identified weaknesses using the Bayesian approach, we compared the predictions of the Bayesian models to the original non-Bayesian models (supplementary Fig. 1<!---\@ref(fig:res-p-pred-diff-hol-intensity2)--->).  
All analyzes were performed in R (`r R.Version()$version.string %>% stringr::str_remove("^R ") %>% stringr::str_remove("\\)") %>% stringr::str_replace(" \\(", ", ")`) [@RCoreTeam.2020]. Spectra were preprocessed using ir (`r packageVersion("ir")`) [@Teickner.2020]. The original script used by @Hodgkins.2018 to extract peaks (\url{https://github.com/shodgkins/FTIRbaselines}) as implemented in irpeat (`r packageVersion("irpeat")`) [@Teickner.2020a] was used to recompute the original models and extract peaks from MIRS. Bayesian models were computed using rstanarm (`r packageVersion("rstanarm")`) [@Goodrich.2020] and brms (`r packageVersion("brms")`) [@Burkner.2017; @Burkner.2018] which are interfaces to Stan (`r rstan::stan_version()`) [@StanDevelopmentTeam.2020; @StanDevelopmentTeam.2021]. All predictor variables were $z$-transformed. MCMC sampling was validated in addition using bayesplot (`r packageVersion("bayesplot")`) [@Gabry.2020]. The out-of-sample predictive performance of the models was estimated using the expected log predictive density (ELPD) estimated using Pareto smoothed importance sampling-leave-one-out cross-validation (PSIS-LOO) [@Vehtari.2017] using the loo package (`r packageVersion("loo")`) [@Vehtari.2019]. PSIS-LOO ELPD is an universal measure to compare the predictive performance of models and $-2 \cdot \text{PSIS-LOO ELPD}$ is the deviance [@Vehtari.2017].

@@

Above, we mentioned that assuming a model with normal distribution can result in unrealistic predictions outside the interval [0, 100] mass-%. We were interested if and under which conditions unrealistic predictions can occur. Here, we differentiate two ways in which a prediction can be unrealistic: It can be either an unrealistic point estimate (median predictions), or the prediction interval can cover values outside [0, 100] mass-%, even though the median prediction can be in [0, 100] mass-%.  
For our analysis, we first computed a sequence of normalized `carb` and `arom15arom16` peak heights such that the original model predict values completely covering the interval [0, 100] mass-% for holocellulose and Klason lignin, respectively. For these simulated peak heights, we computed median model predictions and 90% prediction intervals using the original models ("original Gaussian models"). Finally, we identified unrealistic predictions in terms of the point estimates and 90% prediction intervals, and the `carb` and `arom15arom16` peak heights under which these occur. We used these values as thresholds to decide if the models make unrealistic predictions in practice. To relate these values to real data, we computed median predictions and 90% prediction intervals also for the peat and vegetation samples and identified unrealistic predictions as for the simulation analysis.  
After identifying conditions under which unrealistic predictions occur, we compared this to the behavior of a model with assumptions in line with the data generating process. Beta regression models assume all values of the dependent variable to be in (0,100) mass-% which is a reasonable assumption given how the actual data is generated. Therefore, we repeated the previous analyses, but now using beta regression models. We used the same model structure and priors for intercept and slopes as the original models ("original beta models").  
To facilitate practical comparison between the modeling approaches, we compared the predictions of all models in the form of depth profiles of the holocellulose and Klason lignin contents. Altogether, this allowed us to identify unrealistic predictions of the original models, and to analyze how the improved models (beta regression models) perform in comparison.

@@

1. We used all peak heights and trough heights returned by the peak identification algorithm of @Hodgkins.2018 (Fig. \@ref(fig:hodgkins-peaks-sample-fig01)). The advantage is that this makes the prediction model probably more robust against calibration transfer issues in comparison to the second approach (a hypothesis that remains to be tested). The disadvantage is that additional information contained within the spectra cannot be fully exploited if these are not contained within the extracted peak heights. To avoid overfitting, we computed models using priors implying different amount of regularization for model coefficients (standard deviation of 1 and 0.5, respectively) in addition to the default weakly informative prior which we derived from rstanarm (`r packageVersion("rstanarm")`) (standard deviation of 2.5) [@Goodrich.2020].

2. We used all variables in the spectra and Bayesian regularization [@Piironen.2020], similar to [@Teickner.2022]. The advantage is that this approach can more fully exploit the information contained within the spectra because it is not restricted to specific peaks, resulting potentially in a better predictive performance. The disadvantage is that this approach may be more prone to calibration transfer issues. To reduce redundancy in the variables and impact of noise, we binned the spectra, testing bin widths of 10, 20, 30, 50, and 100 cm$^{-1}$. As regularizing priors, we used the regularized horseshoe prior with a parameterization as described in @Piironen.2017b, assuming a number of relevant variables of eight.

Out-of-sample model performance was estimated with PSIS-LOO ELPD as described above [@Vehtari.2017; @Vehtari.2019]. If a model has a larger PSIS-LOO ELPD than another model fitted on the same data, it has a larger (leave-one-out cross-validation) predictive accuracy, so models with larger PSIS-LOO ELPD are preferred [@Vehtari.2017]. To compare models, one typically computes the posterior distribution of the difference in PSIS-LOO ELPD between the model with the largest PSIS-LOO ELPD and the other models, such that one can evaluate the probability of a certain predictive advantage of the best model in comparison to the other models [@Vehtari.2017; @Vehtari.2019].  
To further validate and interpret the models using binned spectra, we visually identified the most important bins in the models (largest absolute median coefficients $\ge0.2$) with bin a width of 20 cm$^{-1}$ (which are among the models with the best predictive accuracy; see below) and linked these to molecular structures and causal relations that most probably result in the correlation with holocellulose and Klason lignin, respectively.  
Based on the model validation, we defined the following set of models used in the subsequent analyses: "Best all peaks" denotes the models for holocellulose and Klason lignin with best average predictive accuracy of approach 1 described above. "Best binned spectra" denotes the models for holocellulose and Klason lignin with nearly the best average predictive accuracy of approach 2 described above. With "nearly the best predictive accuracy" we mean here that we used the models using binned spectra with a bin width of 20 cm$^{-1}$. This was the model with best average predictive accuracy for Klason lignin, but not holocellulose (Sect. 3.2). However, the model for holocellulose had a predictive performance very similar to the on average better model using a bin width of 10 cm$^{-1}$ (Sect. 3.2), and to keep the following analyses compact and easier to follow, we decided to use a bin width of 20 cm$^{-1}$ in all cases. These models are the models used during the subsequent analyses.

>>>>

    To understand global soil organic matter (SOM) chemistry and its dynamics, we need tools to efficiently quantify SOM properties, for example prediction models using mid infrared spectra. However, the advantages of such models rely on their validity and accuracy. Recently, \cite{Hodgkins.2018} developed models to quantitatively predict peat holocellulose and Klason lignin contents, two indicators of SOM stability and major fractions of organic matter. The models may help to understand large-scale SOM gradients and have been used in various studies.  

@@

    The major limitation we identified is that the original training data are not representative for a range of diverse peat samples. This causes both biased estimates and extrapolation uncertainty under the original models. In addition, the original models can in practice produce unrealistic predictions (negative values or values $>100$ mass-%). Our improved models partly reduce the observed bias, have a better predictive performance for the training data, and avoid such unrealistic predictions. Finally, we provide a proof-of-concept that holocellulose contents can also be predicted for mineral-rich samples (e.g. peat with mineral admixtures or potentially mineral soils).  

@@

Soil and plant OM is often characterized by step-wise chemical fractionation into holocellulose (acid soluble) --- a proxy for polysaccharides --- and lignin (acid insoluble) --- a proxy for aromatics --- [@DelaCruz.2016; @Elle.2019] and these variables are often important indicators for the chemical quality of OM in studies and models analyzing decomposition of SOM [@Leifeld.2012; @Biester.2014; @Worrall.2017; @Hodgkins.2018; @Bengtsson.2018; @Agren.1996a; @Bauer.2004; @Shipley.2021]. If these fractions can be predicted from MIRS, it may be possible to understand decomposition across larger scales and at higher spatial resolution.  
Several models to predict holocellulose and lignin contents in different OM types have been developed (some based on near infrared spectra) [summarized for lignin by @Elle.2019; for holocellulose see e.g. @Peltre.2011; @Sun.2011]. Most of these models consider only wood, material from few woody and non-woody species, or only specific vegetation organs [@Elle.2019]. Another problem is that most of these models cannot be easily reproduced and used since the raw data and code have not been published.  
Recently, @Hodgkins.2018 developed models to predict Klason lignin and holocellulose contents from attenuated total reflectance MIRS. The models have several advantages over existing approaches. The training data comprise several different OM types, such as paper products (cardboard, office paper, magazines, newspaper), leaves from diverse species (trees and graminoids), and wood samples [@DelaCruz.2016; @Hodgkins.2018]. Moreover, both raw data and model code are publicly available [@Hodgkins.2018; @Teickner.2020a]. For this reason, the models are particularly suitable for application in other studies [@Teickner.2019; @Moore.2019; @Harris.2020; @Cong.2020; @Cong.2022], most recently in @Verbeke.2022 and @Baysinger.2022, and for future developments, e.g. by including additional data.   
A major problem is that neither @Hodgkins.2018, nor any later study which was cited above and which used the models provided a thorough validation of the models. To compute the models, @Hodgkins.2018 developed a procedure to extract peaks which are indicative for specific OM fractions, such as holocellulose and Klason lignin (Fig. \@ref(fig:hodgkins-peaks-sample-fig01)), from the MIRS. In the procedure, the peaks are baseline corrected, their maximum height computed, and normalized --- divided by the sum of absorbance values in the spectra. In @Hodgkins.2018, these values were used to calibrate the prediction models. As model validation, only the linearity between the target variables and predictors was assessed (supplementary material to @Hodgkins.2018), but not, for example, how predicted values match measured values. Moreover, no information has been provided whether the training data are representative for SOM, particularly the peat samples analyzed in the study. Furthermore, it is unclear if pre-selecting peaks may reduce the predictive accuracy of the models. Many of these concerns have also been raised by an anonymous reviewer of the paper (supplementary information to @Hodgkins.2018). Thus, an important research gap to fill is to validate the models and to provide concrete strategies for further improvements.

@@

Our goals are to identify limitations in the original models and to give concrete recommendations for improvements. Moreover, we use the original data from @Hodgkins.2018 to provide improvements on the original models where possible. To this end, we conducted an exploratory analysis using the peat and peat-forming vegetation data provided by @Hodgkins.2018. Our exploratory analysis was guided by the following research questions:

1. _Is the normal distribution reasonable to predict holocellulose and Klason lignin contents for peat samples?_ Klason lignin and holocellulose contents as actually measured for the training data set in @Hodgkins.2018 cannot be negative or larger than 100 mass-%. For this reason, predictions and prediction uncertainties of the models should be in the interval [0, 100] mass-%. Yet, the original models assume a normal distribution which does not entail any lower or upper bound for predicted values. This can potentially result in unrealistic predictions. However, if no unrealistic predictions occur for "representative" peat samples, using a model with normal distribution can be justifiable in practice. In contrast, if unrealistic predictions occur, it is more reasonable to use a distribution for which assumptions are consistent with knowledge on the data generating process. For example, the beta distribution assumes a lower and upper bound which can be mapped to the interval (0, 100) mass-%. It can be used as replacement for the normal sampling distribution used in ordinary linear regression models, such as the original models. For this reason, we were interested if and under which conditions the original models may produce unrealistic predictions for the peat and vegetation samples.

@@

3. _Is the prediction domain of the training data representative for peat samples?_ If a model extrapolates outside the prediction domain --- the range of predictor variable values in the training data set [@Wadoux.2021] --- it is unclear if the predictions are valid. An advantage of the independent reference sample set used by @Hodgkins.2018 is that the model is applicable to various OM types and, for example, does not depend on obviously material-specific spectral characteristics [@Hodgkins.2018]. For example, in peat, holocellulose content is controlled by decomposition, and other structures also controlled by decomposition might therefore provide information on holocellulose contents [@Leifeld.2012; @Biester.2014]. However, this does not have to be the case for other samples, e.g. peat forming vegetation. A potential advantage is therefore the potential generality of the training data set @Hodgkins.2018 used. However, since it does not include peat samples, it is questionable if the prediction domain formed by the training samples covers peat samples and peat forming vegetation in general, and this can result in overlooking prediction failures due to extrapolation. For this reason, we were interested if the training data (its prediction domain, @Wadoux.2021) is representative for peat and peat forming vegetation.

4. _Do predictions of the improved models differ and if so, why?_ Computing modified models enables us to compare their predictions to those of the original models using the same training data, and to the actually measured contents. Model comparison often reveals detailed insights into the mechanics of a model and may provide hints to which model makes more correct predictions. We therefore were interested in analyzing differences between the computed models, what factors may cause differences in predictions, and if there are indications which models make more correct predictions.

5. _Can holocellulose contents also be predicted for samples with mineral interferences?_ During model calibration, @Hodgkins.2018 omitted some samples. For prediction of holocellulose, old magazine samples were omitted because the `carb` peak suffered from mineral interference from the samples' clay coatings [@Hodgkins.2018]. This decision is sensible, but since there are many organic soils with mineral admixtures --- for example due to volcanic ash, in base layers, or due to cryoturbation) [e.g. @Broder.2012; @Bockheim.2007; @Koven.2009; @Loisel.2014] --- we think that it is useful to have a model to predict holocellulose contents that includes mineral-rich samples. We therefore wanted to investigate if a suitable model can be computed by including the previously omitted old magazine samples during calibration. This is no replacement for a model using more training data, but it is a test whether it is at least in principle possible for MIRS to contain sufficient information to predict holocellulose contents of relatively diverse samples.

@@

We conducted a series of statistical analyzes to query each of our research questions listed in the introduction. During this, we also computed improved models using the same training data that was used to compute the original models. Since we used these improved models to uncover and analyze the limitations in the original models, we first describe steps how we improved the original models and then how we analyzed the limitations.  
We use the data and models collected and developed by @Hodgkins.2018. The training data used to calibrate the original models consist of various organic materials and comprise measured values for holocellulose and Klason lignin [@DelaCruz.2016]. The data are accessible from the supplementary information of @Hodgkins.2018. In addition, we used a data set consisting of MIRS for `r length(unique(d_peat$core_label))` peat cores (`r nrow(d_peat)` samples in total) and `r nrow(d_veg)` vegetation samples collected at the same sites which are also provided by @Hodgkins.2018. We used these peat and vegetation data to evaluate the performance of the original and improved models for peat and peat forming vegetation in general. We assume that the peat samples are representative for a wide range of peat OM because they cover a large latitudinal range and a wide range of degrees of decomposition, and were formed by different vegetation [@Hodgkins.2018].  
We computed all models as Bayesian models. This allowed us to consider parameter uncertainty in predictions and to compute models with good predictive performance [@Raftery.2003; @Piironen.2017]. For this reason, to facilitate model comparison, we also recomputed the original models as Bayesian models with weakly informative priors. With weakly informative priors, we mean here priors which result in approximately the same prediction intervals as the original frequentist models (supplementary Fig. 1<!---\@ref(fig:res-p-pred-diff-hol-intensity2)--->). Whenever we refer to "original model", we mean these Bayesian translations, and otherwise refer to "original non-Bayesian models". To check that we did not introduce any of the identified weaknesses using the Bayesian approach, we compared the predictions of the Bayesian models to the original non-Bayesian models (supplementary Fig. 1<!---\@ref(fig:res-p-pred-diff-hol-intensity2)--->).  
All analyzes were performed in R (`r R.Version()$version.string %>% stringr::str_remove("^R ") %>% stringr::str_remove("\\)") %>% stringr::str_replace(" \\(", ", ")`) [@RCoreTeam.2020]. Spectra were preprocessed using ir (`r packageVersion("ir")`) [@Teickner.2020]. The original script used by @Hodgkins.2018 to extract peaks (\url{https://github.com/shodgkins/FTIRbaselines}) as implemented in irpeat (`r packageVersion("irpeat")`) [@Teickner.2020a] was used to recompute the original models and extract peaks from MIRS. Bayesian models were computed using rstanarm (`r packageVersion("rstanarm")`) [@Goodrich.2020] and brms (`r packageVersion("brms")`) [@Burkner.2017; @Burkner.2018] which are interfaces to Stan (`r rstan::stan_version()`) [@StanDevelopmentTeam.2020; @StanDevelopmentTeam.2021]. All predictor variables were $z$-transformed. Markov Chain Monte Carlo (MCMC) sampling was validated in addition using bayesplot (`r packageVersion("bayesplot")`) [@Gabry.2020]. The out-of-sample predictive performance of the models was estimated using the expected log predictive density (ELPD) estimated using Pareto smoothed importance sampling-leave-one-out cross-validation (PSIS-LOO) [@Vehtari.2017] using the loo package (`r packageVersion("loo")`) [@Vehtari.2019]. PSIS-LOO ELPD is an universal measure to compare the predictive performance of models and $-2 \cdot \text{PSIS-LOO ELPD}$ is the deviance [@Vehtari.2017].

@@

Above, we mentioned that assuming a model with normal distribution can result in unrealistic predictions outside the interval [0, 100] mass-%. We were interested if and under which conditions unrealistic predictions can occur. Here, we differentiate two ways in which a prediction can be unrealistic: It can be either an unrealistic point estimate (median predictions), or the prediction interval can cover values outside [0, 100] mass-%, even though the median prediction can be within [0, 100] mass-%.  
For our analysis, we first computed a sequence of normalized `carb` and `arom15arom16` peak heights such that the original model predict values which completely covered the interval [0, 100] mass-% for holocellulose and Klason lignin, respectively. For these simulated peak heights, we computed median model predictions and 90% prediction intervals using the original models ("original Gaussian models"). Finally, we identified unrealistic predictions in terms of the point estimates and 90% prediction intervals, and the `carb` and `arom15arom16` peak heights under which these occur. We used these values as thresholds to decide if the models make unrealistic predictions in practice. To relate these values to real data, we computed median predictions and 90% prediction intervals also for the peat and vegetation samples and identified unrealistic predictions as for the simulation analysis.  
After identifying conditions under which unrealistic predictions occur, we compared this to the behavior of a model with assumptions in line with the data generating process. Beta regression models assume all values of the dependent variable to be in (0,100) mass-% which is a reasonable assumption given how the actual data are generated. Therefore, we repeated the previous analyses, but now using beta regression models. We used the same model structure and priors for intercept and slopes as the original models ("original beta models").  
To facilitate practical comparison between the modeling approaches, we compared the predictions of all models for the peat samples by plotting predicted values versus depths of the peat samples. Altogether, this allowed us to identify unrealistic predictions of the original models, and to analyze how the improved models (beta regression models) perform in comparison.

@@

1. We used all peak heights and trough heights returned by the peak identification algorithm of @Hodgkins.2018 (Fig. \@ref(fig:hodgkins-peaks-sample-fig01)). The advantage is that this makes the prediction model probably more robust against calibration transfer issues in comparison to the second approach (a hypothesis that remains to be tested). The disadvantage is that additional information contained within the spectra cannot be fully exploited if these are not contained within the extracted peak heights. To avoid overfitting, we computed models using priors implying different amounts of regularization --- shrinking coefficients to 0 with the aim to avoid overfitting --- of model coefficients (standard deviation of 1 and 0.5, respectively) in addition to the default weakly informative prior which we derived from rstanarm (`r packageVersion("rstanarm")`) (standard deviation of 2.5) [@Goodrich.2020].

2. We used all variables in the spectra and Bayesian regularization [@Piironen.2020], similar to @Teickner.2022. The advantage is that this approach can more fully exploit the information contained within the spectra because it is not restricted to specific peaks, resulting potentially in a better predictive performance. The disadvantage is that this approach may be more prone to calibration transfer issues. To reduce redundancy in the variables and impact of noise, we binned the spectra, testing bin widths of 10, 20, 30, 50, and 100 cm$^{-1}$. As regularizing priors, we used the regularized horseshoe prior with a parameterization as described in @Piironen.2017b, assuming a number of relevant variables of eight.

### 2022-07-25 analysis/paper/002-reply-main.Rmd
Address reviewer comments to manuscript 1.

---
title: 'Predicting Absolute Holocellulose and Klason Lignin Contents for Peat Remains Challenging'
author: "Henning Teickner$^{1, *}$, Klaus-Holger Knorr$^1$"
date: ""
output: 
  bookdown::pdf_book:
    keep_tex: true
    citation_package: natbib
    toc: false
bibliography: references
biblio-style: unsrtnat
natbiboptions: super
header-includes:
   - \usepackage{float}
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
---

>>>>

---
title: 'Comment on Hodgkins et al. (2018): Predicting Absolute Holocellulose and Klason Lignin Contents for Peat Remains Challenging'
author: "Henning Teickner$^{1, *}$, Klaus-Holger Knorr$^1$"
date: ""
output: 
  bookdown::pdf_book:
    keep_tex: true
    citation_package: natbib
    toc: false
bibliography: references
biblio-style: unsrtnat
natbiboptions: super
header-includes:
   - \usepackage{float}
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
---
